---
title: "MATH 216 Homework 3"
author: "Alexander Pastora"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)
library(Quandl)
library(forcats)
```



## Admistrative:

Please indicate

* Who you collaborated with: None
* Roughly how much time you spent on this HW so far: 9
* The URL of the RPubs published URL [here](http://rpubs.com/pinkfan9999/HW-3).
* What gave you the most trouble:
* Any comments you have:




## Question 1:

We will use a logistic regression model to predict sex. Our metric to rate how well our
model performs will be:

$$
\frac{1}{n}\sum_{i=1}^{n}I(y_i = \widehat{y}_i)
$$

where $I(A)$ is the *indicator function* that is equal to 1 if condition $A$
holds, 0 otherwise. So

* Say user $i$ **is** female, then $y_i=1$
* Say we **predict** user $i$ is female, then $\widehat{y}_i=1$
* In this case $I(y_i =\widehat{y}_i)=1$. 

So what the above formula is reporting is the proportion of users' sex we
correctly predicted.

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
# To keep this exercise simple, let's remove the only 3 users (0.005% of users)
# who did not list a height, define the outcome variable, and add an ID variable
# to distinguish the users
profiles <- read_csv(file="profiles.csv") %>% 
  filter(!is.na(height)) %>% 
  mutate(is_female=ifelse(sex=='f', 1, 0)) %>% 
  tibble::rownames_to_column(var="id")
```



#### a)

Define:

* A *training* set `training` of 2997 users (5% of users). We will train the 
logistic regression model to predict gender using this data. Since we want to 
train the model to tell who is female and who is not, we use the outcome
variable `is_female`.
* A *test* set `test` of the remaining 56,946 users (95% of users). We will test
how good our trained model is using this data. So at first, we will pretend we
don't know the outcome variable `is_female`. We use the above model to make a
prediction of sex for all 56,946 test users, then we use the `is_female` outcome
to rate how well we performed.
* Be sure to incorporate all the insight your garnered in your EDA in HW-2.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
set.seed(76)
#Consolidate categories + rename NA for job data
profiles <-profiles %>% mutate(job = fct_explicit_na(job,na_level = "Not Reported")) %>% mutate(job =fct_recode(job,
                                                                                                        "professional"="law / legal services",                                                                                               
                                                                                                        "professional" = "medicine / health",
                                                                                                        "entertainment" = "artistic / musical / writer",
                                                                                                        "entertainment" = "entertainment / media",
                                                                                                        "administration" = "clerical / administrative",
                                                                                                        "administration" = "executive / management",
                                                                                                        "finance" = "banking / financial / real estate",
                                                                                                        "finance" = "sales / marketing / biz dev",
                                                                                                        "government" = "political / government",
                                                                                                        "government" = "military",
                                                                                                        "tech" = "computer / hardware / software",
                                                                                                        "tech" = "science / tech / engineering",
                                                                                                        "travel" = "hospitality / travel",
                                                                                                        "travel" = "transportation",
                                                                                                        "none" = "retired",
                                                                                                        "none" = "unemployed",
                                                                                
                                                                                                        "other" = "other",
                                                                                                        "other" = "rather not say",
                                                                                                        "education" = "education / academia",
                                                                                                        "construction" = "construction / craftsmanship")
                                                                                                        )

# Consolidate categories + rename NA for diet
profiles <- profiles %>% mutate(diet = fct_explicit_na(diet,na_level = "Not Reported")) %>%  mutate(diet = fct_recode(diet,"anything" = "anything", "religious diet/other"="halal", "religious diet/other"="kosher", "anything"="mostly anything" , "religious diet/other"="mostly halal" , "religious diet/other"="mostly kosher","religious diet/other"= "mostly other","vegetarian/vegan"= "mostly vegan", "vegetarian/vegan"="mostly vegetarian", "religious diet/other"="other", "anything" = "strictly anything", "religious diet/other"="strictly halal", "religious diet/other"="strictly kosher", "religious diet/other"="strictly other" , "vegetarian/vegan"="strictly vegan" , "vegetarian/vegan"="strictly vegetarian" , "vegetarian/vegan"="vegan", "vegetarian/vegan"="vegetarian"))

# Create training and test sets
training <- profiles %>% sample_n(2997)
test <- anti_join(profiles, training, by="id")

```



#### b)

Train the logistic regression model to predict sex. i.e. fit a logistic
regression model to the `training` data. Assign this model to an R object called
`predict_sex_model`, then rate how well the model performs on the `training` data.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
predict_sex_model <- glm(is_female ~ diet + job, data=training, family="binomial")
broom::tidy(predict_sex_model)


```



#### c)

Take `predict_sex_model` and apply it to the `test` data and make a prediction 
for each users' sex, then rate how well the model performs on the `test` data.

**Hint**: What do you think `predict(predict_sex_model, newdata=test,
type="response")` does? The help file is located in `?predict.glm`

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
p_hat <- predict(predict_sex_model, newdata=test, type="response")
ggplot(data=NULL, aes(x=p_hat)) + 
  geom_histogram()

test <- test %>% 
  mutate(p_hat = predict(predict_sex_model, newdata=test, type="response"))

test %>% 
  select(job, diet, p_hat, is_female)
```

Overall, the model performed decently well on the test data. However, when the p-hat values were around .45-.65, the predictions got a little inconsistent. 


#### d)

Did the model perform better on the `training` data or the `test` data? Why
do you think that is?

The model definitely peformed better on the training data, because the model was based off of the training data. 





## Question 2:

We want to compare the volatility of 

* [Bitcoin](https://www.quandl.com/data/BAVERAGE/USD) prices
* [Gold](https://www.quandl.com/data/BUNDESBANK/BBK01_WT5511) prices

Let our measure of volatility be the relative change from day-to-day in price. 
Let the reference currency be US dollars. Analyze these results and provide
insight to a foreign currency exchanger.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
bitcoin <- Quandl("BAVERAGE/USD") %>%
  tbl_df() %>%
  rename(
    Avg = `24h Average`,
    Total_Volume = `Total Volume`
  )
ggplot(data=bitcoin, aes(x=Date, y=Avg)) +geom_point()
bitcoin_changes <- bitcoin %>%  mutate(previous_day = lead(Date)) 
bitcoin_data <- full_join(bitcoin_changes, bitcoin, by= c("previous_day"= "Date"))  %>% mutate(percent_change = ((Avg.x-Avg.y)/Avg.y)*100) %>% filter(!is.na(percent_change))



gold <- Quandl("BUNDESBANK/BBK01_WT5511") %>%
  tbl_df() 
ggplot(data=gold, aes(x=Date, y=Value)) +geom_point()
gold_changes <- gold %>% mutate(previous_date = lead(Date))
gold_data <- full_join(gold_changes, gold, by= c("previous_date"= "Date")) %>% mutate(percent_change = ((Value.x-Value.y)/Value.y)*100) %>% filter(!is.na(percent_change)) 

gold_data <- gold_data %>% mutate(Type="Gold") %>% select(Date, percent_change, Type)
bitcoin_data <- bitcoin_data %>% mutate(Type="bitcoin") %>% select(Date, percent_change, Type)
combined_data <- bind_rows(gold_data, bitcoin_data)
ggplot(combined_data, aes(x=Date, y=percent_change, col=Type)) +
geom_line() + coord_cartesian(xlim=c(ymd(20100718),ymd(20160418))) + ylab("Percent Change") + ggtitle("Percent Change of Gold and Bitcoin")

```

Bitcoin is more volatile than Gold, because Bitcoin has larger relative changes per day than Gold. 




## Question 3:

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
jukebox <- read_csv(file="reed_jukebox.csv")

# Clean certain artists' names:
sigur_ros <- "Sigur Ro\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8c\x93\xa0\xbcs"
bjork <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk"
blue_oyster_cult <- "Blue O\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcyster Cult"
husker_do <- "Hu\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcsker Du\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbc"
bjork_brodsky <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk & Brodsky Quartet"
slagsmalsklubben <- "Slagsma\xfc\xbe\x99\x86\x94\xbc_lsklubben "
bjork_sugarcubes <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk (Sugarcubes)"

jukebox <- jukebox %>%
  mutate(
    artist = ifelse(artist == sigur_ros, "Sigor Ros", artist),
    artist = ifelse(artist == bjork, "Bjork", artist),
    artist = ifelse(artist == blue_oyster_cult, "Blue Oyster Cult", artist),
    artist = ifelse(artist == husker_do, "Husker Do", artist),
    artist = ifelse(artist == bjork_brodsky, "Bjork & Brodsky Quartet", artist),
    artist = ifelse(artist == slagsmalsklubben, "Slagsmalsklubben", artist),
    artist = ifelse(artist == bjork_sugarcubes, "Bjork (Sugarcubes)", artist)
  )
```


Using the Reed College jukebox data, what are the top 10 artists played during
the "graveyard shift" during the academic year? Define

* the "graveyard shift" as midnight to 8am
* the academic year as September through May (inclusive)

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
academic_year <- jukebox %>% mutate(date_time=parse_date_time(date_time,"a b! d! H: M:S Y!"))
academic_year <- academic_year %>% mutate(month= month(date_time)) %>% filter(month <= 5 | month >=9)


graveyard <- academic_year %>% mutate(hour = hour(date_time)) %>% filter(hour>= 0 & hour <= 8) 
graveyard_artist <- graveyard %>% group_by(artist) %>% tally()
graveyard_artist %>% top_n(10) %>% knitr::kable()


```





